### 2024-11-28 joss/paper.preprint.tex
update heuristic parameter convergence plot, add tex preprint format.

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% definitions for citeproc citations
\NewDocumentCommand\citeproctext{}{}
\NewDocumentCommand\citeproc{mm}{%
  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
\makeatletter
 % allow citations to break across lines
 \let\@cite@ofmt\@firstofone
 % avoid brackets around text for \cite:
 \def\@biblabel#1{}
 \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
 {\begin{list}{}{%
  \setlength{\itemindent}{0pt}
  \setlength{\leftmargin}{0pt}
  \setlength{\parsep}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
   \setlength{\leftmargin}{\cslhangindent}
   \setlength{\itemindent}{-1\cslhangindent}
  \fi
  % set entry spacing
  \setlength{\itemsep}{#2\baselineskip}}}
 {\end{list}}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\ifLuaTeX
\usepackage[bidi=basic]{babel}
\else
\usepackage[bidi=default]{babel}
\fi
\babelprovide[main,import]{american}
% get rid of language-specific shorthands (see #6817):
\let\LanguageShortHands\languageshorthands
\def\languageshorthands#1{}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={PyBOP: A Python package for battery model optimisation and
parameterisation},
  pdfauthor={Brady Planden, Nicola E. Courtier, Martin Robinson, Ferran
Brosa Planella, David A. Howey},
  pdflang={en-US},
  colorlinks=true,
  linkcolor={Maroon},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{PyBOP: A Python package for battery model optimisation and
parameterisation}

\definecolor{c53baa1}{RGB}{83,186,161}
\definecolor{c202826}{RGB}{32,40,38}
\def \rorglobalscale {0.1}
\newcommand{\rorlogo}{%
\begin{tikzpicture}[y=1cm, x=1cm, yscale=\rorglobalscale,xscale=\rorglobalscale, every node/.append style={scale=\rorglobalscale}, inner sep=0pt, outer sep=0pt]
  \begin{scope}[even odd rule,line join=round,miter limit=2.0,shift={(0,3.0889892500000005)},xscale=0.9943,yscale=0.9894]
    \path[fill=white,even odd rule,line join=round,miter limit=2.0] (0.0, 3.1221) rectangle (4.364, -0.0336);
  \end{scope}
  \begin{scope}[even odd rule,line join=round,miter limit=2.0,shift={(-0.025, 0.0216)}]
    \path[fill=c53baa1,nonzero rule,line join=round,miter limit=2.0] (1.8164, 3.012) -- (1.4954, 2.5204) -- (1.1742, 3.012) -- (1.8164, 3.012) -- cycle;
    \path[fill=c53baa1,nonzero rule,line join=round,miter limit=2.0] (3.1594, 3.012) -- (2.8385, 2.5204) -- (2.5172, 3.012) -- (3.1594, 3.012) -- cycle;
    \path[fill=c53baa1,nonzero rule,line join=round,miter limit=2.0] (1.1742, 0.0669) -- (1.4954, 0.5588) -- (1.8164, 0.0669) -- (1.1742, 0.0669) -- cycle;
    \path[fill=c53baa1,nonzero rule,line join=round,miter limit=2.0] (2.5172, 0.0669) -- (2.8385, 0.5588) -- (3.1594, 0.0669) -- (2.5172, 0.0669) -- cycle;
    \path[fill=c202826,nonzero rule,line join=round,miter limit=2.0] (3.8505, 1.4364).. controls (3.9643, 1.4576) and (4.0508, 1.5081) .. (4.1098, 1.5878).. controls (4.169, 1.6674) and (4.1984, 1.7642) .. (4.1984, 1.8777).. controls (4.1984, 1.9719) and (4.182, 2.0503) .. (4.1495, 2.1132).. controls (4.1169, 2.1762) and (4.0727, 2.2262) .. (4.0174, 2.2635).. controls (3.9621, 2.3006) and (3.8976, 2.3273) .. (3.824, 2.3432).. controls (3.7505, 2.359) and (3.6727, 2.367) .. (3.5909, 2.367) -- (2.9676, 2.367) -- (2.9676, 1.8688).. controls (2.9625, 1.8833) and (2.9572, 1.8976) .. (2.9514, 1.9119).. controls (2.9083, 2.0164) and (2.848, 2.1056) .. (2.7705, 2.1791).. controls (2.6929, 2.2527) and (2.6014, 2.3093) .. (2.495, 2.3487).. controls (2.3889, 2.3881) and (2.2728, 2.408) .. (2.1468, 2.408).. controls (2.0209, 2.408) and (1.905, 2.3881) .. (1.7986, 2.3487).. controls (1.6925, 2.3093) and (1.6007, 2.2527) .. (1.5232, 2.1791).. controls (1.4539, 2.1132) and (1.3983, 2.0346) .. (1.3565, 1.9436).. controls (1.3504, 2.009) and (1.3351, 2.0656) .. (1.3105, 2.1132).. controls (1.2779, 2.1762) and (1.2338, 2.2262) .. (1.1785, 2.2635).. controls (1.1232, 2.3006) and (1.0586, 2.3273) .. (0.985, 2.3432).. controls (0.9115, 2.359) and (0.8337, 2.367) .. (0.7519, 2.367) -- (0.1289, 2.367) -- (0.1289, 0.7562) -- (0.4837, 0.7562) -- (0.4837, 1.4002) -- (0.6588, 1.4002) -- (0.9956, 0.7562) -- (1.4211, 0.7562) -- (1.0118, 1.4364).. controls (1.1255, 1.4576) and (1.2121, 1.5081) .. (1.2711, 1.5878).. controls (1.2737, 1.5915) and (1.2761, 1.5954) .. (1.2787, 1.5991).. controls (1.2782, 1.5867) and (1.2779, 1.5743) .. (1.2779, 1.5616).. controls (1.2779, 1.4327) and (1.2996, 1.3158) .. (1.3428, 1.2113).. controls (1.3859, 1.1068) and (1.4462, 1.0176) .. (1.5237, 0.944).. controls (1.601, 0.8705) and (1.6928, 0.8139) .. (1.7992, 0.7744).. controls (1.9053, 0.735) and (2.0214, 0.7152) .. (2.1474, 0.7152).. controls (2.2733, 0.7152) and (2.3892, 0.735) .. (2.4956, 0.7744).. controls (2.6016, 0.8139) and (2.6935, 0.8705) .. (2.771, 0.944).. controls (2.8482, 1.0176) and (2.9086, 1.1068) .. (2.952, 1.2113).. controls (2.9578, 1.2253) and (2.9631, 1.2398) .. (2.9681, 1.2544) -- (2.9681, 0.7562) -- (3.3229, 0.7562) -- (3.3229, 1.4002) -- (3.4981, 1.4002) -- (3.8349, 0.7562) -- (4.2603, 0.7562) -- (3.8505, 1.4364) -- cycle(0.9628, 1.7777).. controls (0.9438, 1.7534) and (0.92, 1.7357) .. (0.8911, 1.7243).. controls (0.8623, 1.7129) and (0.83, 1.706) .. (0.7945, 1.7039).. controls (0.7588, 1.7015) and (0.7252, 1.7005) .. (0.6932, 1.7005) -- (0.4839, 1.7005) -- (0.4839, 2.0667) -- (0.716, 2.0667).. controls (0.7477, 2.0667) and (0.7805, 2.0643) .. (0.8139, 2.0598).. controls (0.8472, 2.0553) and (0.8768, 2.0466) .. (0.9025, 2.0336).. controls (0.9282, 2.0206) and (0.9496, 2.0021) .. (0.9663, 1.9778).. controls (0.9829, 1.9534) and (0.9914, 1.9209) .. (0.9914, 1.8799).. controls (0.9914, 1.8362) and (0.9819, 1.8021) .. (0.9628, 1.7777) -- cycle(2.6125, 1.3533).. controls (2.5889, 1.2904) and (2.5553, 1.2359) .. (2.5112, 1.1896).. controls (2.4672, 1.1433) and (2.4146, 1.1073) .. (2.3529, 1.0814).. controls (2.2916, 1.0554) and (2.2228, 1.0427) .. (2.1471, 1.0427).. controls (2.0712, 1.0427) and (2.0026, 1.0557) .. (1.9412, 1.0814).. controls (1.8799, 1.107) and (1.8272, 1.1433) .. (1.783, 1.1896).. controls (1.7391, 1.2359) and (1.7052, 1.2904) .. (1.6817, 1.3533).. controls (1.6581, 1.4163) and (1.6465, 1.4856) .. (1.6465, 1.5616).. controls (1.6465, 1.6359) and (1.6581, 1.705) .. (1.6817, 1.7687).. controls (1.7052, 1.8325) and (1.7388, 1.8873) .. (1.783, 1.9336).. controls (1.8269, 1.9799) and (1.8796, 2.0159) .. (1.9412, 2.0418).. controls (2.0026, 2.0675) and (2.0712, 2.0804) .. (2.1471, 2.0804).. controls (2.223, 2.0804) and (2.2916, 2.0675) .. (2.3529, 2.0418).. controls (2.4143, 2.0161) and (2.467, 1.9799) .. (2.5112, 1.9336).. controls (2.5551, 1.8873) and (2.5889, 1.8322) .. (2.6125, 1.7687).. controls (2.636, 1.705) and (2.6477, 1.6359) .. (2.6477, 1.5616).. controls (2.6477, 1.4856) and (2.636, 1.4163) .. (2.6125, 1.3533) -- cycle(3.8015, 1.7777).. controls (3.7825, 1.7534) and (3.7587, 1.7357) .. (3.7298, 1.7243).. controls (3.701, 1.7129) and (3.6687, 1.706) .. (3.6333, 1.7039).. controls (3.5975, 1.7015) and (3.5639, 1.7005) .. (3.5319, 1.7005) -- (3.3226, 1.7005) -- (3.3226, 2.0667) -- (3.5547, 2.0667).. controls (3.5864, 2.0667) and (3.6192, 2.0643) .. (3.6526, 2.0598).. controls (3.6859, 2.0553) and (3.7155, 2.0466) .. (3.7412, 2.0336).. controls (3.7669, 2.0206) and (3.7883, 2.0021) .. (3.805, 1.9778).. controls (3.8216, 1.9534) and (3.8301, 1.9209) .. (3.8301, 1.8799).. controls (3.8301, 1.8362) and (3.8206, 1.8021) .. (3.8015, 1.7777) -- cycle;
  \end{scope}
\end{tikzpicture}
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Authors and Affiliations

\usepackage[affil-it]{authblk}
\usepackage{orcidlink}
%% \renewcommand\Authsep{, }
\setlength{\affilsep}{1em}
\author[1%
  %
  \ensuremath\mathparagraph]{Brady Planden%
    \,\orcidlink{0000-0002-1082-9125}\,%
    }
\author[1,2%
  %
  ]{Nicola E. Courtier%
    \,\orcidlink{0000-0002-5714-1096}\,%
    }
\author[3%
  %
  ]{Martin Robinson%
    \,\orcidlink{0000-0002-1572-6782}\,%
    }
\author[2,4%
  %
  ]{Ferran Brosa Planella%
    \,\orcidlink{0000-0001-6363-2812}\,%
    }
\author[1,2%
  %
  ]{David A. Howey%
    \,\orcidlink{0000-0002-0620-3955}\,%
    }

\affil[1]{Department of Engineering Science, University of Oxford,
Oxford, UK%
  }
\affil[2]{The Faraday Institution, Harwell Campus, Didcot, UK%
  }
\affil[3]{Research Software Engineering Group, University of Oxford,
Oxford, UK%
  }
\affil[4]{Mathematics Institute, University of Warwick, Coventry, UK%
  }
\affil[$\mathparagraph$]{Corresponding author: %
   %
}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\date{28 November 2024}

\begin{document}
\maketitle

\section{Summary}\label{summary}

The Python Battery Optimisation and Parameterisation (\texttt{PyBOP})
package provides a set of methods for the parameterisation and
optimisation of battery models, offering both Bayesian and frequentist
approaches with example workflows to assist the user. \texttt{PyBOP} has
been developed to enable parameter identification of various battery
models, including the electrochemical and equivalent circuit models
provided by the popular open-source \texttt{PyBaMM} package
(\citeproc{ref-Sulzer:2021}{Sulzer et al., 2021}).

Similarly, \texttt{PyBOP} can be used for parameter design optimisation
under user-defined operating conditions across a variety of model
structures. \texttt{PyBOP} allows battery model parameterisation using a
range of methods with diagnostics on the performance and convergence of
the identified or optimised parameters. The identified parameters can be
used for prediction, on-line control and design optimisation, all of
which support improved battery utilisation and development.

\section{Statement of need}\label{statement-of-need}

\texttt{PyBOP} is a Python package designed to provide a user-friendly,
object-oriented interface for optimising battery models. \texttt{PyBOP}
leverages the open-source \texttt{PyBaMM} package
(\citeproc{ref-Sulzer:2021}{Sulzer et al., 2021}) to formulate and solve
of these battery models. \texttt{PyBOP} is intended to serve a broad
audience of students, engineers, and researchers in both academia and
the battery industry by enabling the use of predictive battery models
where previously this was not possible. \texttt{PyBOP} emphasises clear
and informative diagnostics and workflows for users of varying
expertise, by providing advanced optimisation and sampling algorithms.
These methods are provided through interfaces to \texttt{PINTS}
(\citeproc{ref-Clerx:2019}{Clerx et al., 2019}), \texttt{SciPy}
(\citeproc{ref-SciPy:2020}{Virtanen et al., 2020}), in addition to the
PyBOP's own algorithms such as Adaptive Moment Estimation with Weight
Decay (AdamW), and Cuckoo search.

\texttt{PyBOP} supports the Battery Parameter eXchange (BPX) standard
(\citeproc{ref-BPX:2023}{Korotkin et al., 2023}) for sharing battery
parameter sets. As these parameter sets are costly to obtain due to: the
equipment and time required for characterisation experiments, the need
for battery domain knowledge, and the computational cost of parameter
estimation. \texttt{PyBOP} reduces these costs by providing fast
computational estimation with parameter set interoperability.

This package complements other lithium-ion battery modelling packages
built around \texttt{PyBaMM}, such as \texttt{liionpack} for battery
pack simulation (\citeproc{ref-Tranter2022}{Tranter et al., 2022}) and
\texttt{pybamm-eis} for numerical impedance spectroscopy, as the
identified parameters from \texttt{PyBOP} are easily exportable.

\section{Architecture}\label{architecture}

\texttt{PyBOP} has a layered data structure designed to compute and
process the forward model predictions and to package the necessary
information for the optimisation and sampling algorithms. The forward
model is solved using the popular battery modelling package,
\texttt{PyBaMM}, with construction, parameterisation, and discretisation
managed by PyBOP's model interface to PyBaMM. This approach provides a
robust object construction process with a consistent interface between
the models and optimisers. The statistical methods and optimisation
algorithms are then constructed to interface cleanly with the forward
model predictions. Furthermore, identifiability metrics are provided
alongside the estimated parameters through Hessian approximation of the
cost functions in the frequentist workflows and posterior moments in the
Bayesian workflows.

\begin{figure}
\centering
\includegraphics[width=0.8\textwidth,height=\textheight]{figures/PyBOP-high-level.pdf}
\caption{PyBOP's interface to supporting funding agencies, alongside a
visualisation of the general workflow for parameterisation and
optimisation \label{fig:high-level}}
\end{figure}

\texttt{PyBOP} formulates the inference process into four key classes,
namely the model, the problem, the cost, and the optimiser/sampler, as
shown in \autoref{fig:classes}. Each of these objects represents a base
class with child classes constructing specialised functionality for
inference or optimisation workflows. The model class constructs a
\texttt{PyBaMM} forward model for a given set of model equations
provided by \texttt{PyBaMM}, initial conditions, spatial discretisation,
and numerical solver. By composing \texttt{PyBaMM} directly into
\texttt{PyBOP}, specialised models can be constructed alongside the
standard models, which can be modified, and optimally constructed for
the inference tasks. One such example is spatial rediscretisation, which
is performed when geometric parameters are optimised. In this situation,
\texttt{PyBOP} minimally rediscretises the \texttt{PyBaMM} model while
maintaining the problem, cost, and optimiser objects, providing improved
performance benefits to users. Alongside construction of the forward
model, \texttt{PyBOP}`s model class provides methods for obtaining
sensitivities from the prediction, enabling gradient-based optimisation
algorithms. This prediction, along with it's corresponding
sensitivities, is provided to the problem class for processing and
exception control. A standardised data structure is then provided to the
cost classes, providing a distance, design, or likelihood-based metric
for optimisation. For deterministic optimisation, the optimisers
minimise the corresponding cost function or the negative log-likelihood
if a likelihood class is provided. Bayesian inference is provided by
Monte Carlo sampling classes, which accept the LogPosterior class and
sample from it using Pints' based Monte Carlo algorithms at the time of
submission. In the typical workflow, the classes in
\autoref{fig:classes} are constructed in sequence.

In addition to the core architecture, \texttt{PyBOP} provides several
specialised inference and optimisation processes. One such instance is
numerical electrochemical impedance spectroscopy predictions by
discretising the forward model into sparse mass matrix form with
accompanying auto-differentiated Jacobian. These objects are then
translated into the frequency domain with a linear solution used to
compute the battery model impedance. In this situation, the forward
models are constructed within the spatial rediscretisation workflow,
allowing for geometric parameter inference from EIS forward model
predictions. Furthermore, \texttt{PyBOP} builds on the JAX
(\citeproc{ref-jax:2018}{Bradbury et al., 2018}) numerical solvers
provided by \texttt{PyBaMM} by providing JAX-based cost functions for
automatic forward model differentiation with respect to the parameters.
This functionality provides a performance improvement alongside an
interface to JAX-based inference packages, such as Numpyro
(\citeproc{ref-numpyro:2019}{Phan et al., 2019}), BlackJAX
(\citeproc{ref-blackjax:2024}{Cabezas et al., 2024}), and Optax
(\citeproc{ref-optax:2020}{DeepMind et al., 2020}).

\begin{figure}
\centering
\includegraphics[width=0.8\textwidth,height=\textheight]{figures/PyBOP_components.drawio.png}
\caption{The core \texttt{PyBOP} architecture, showcasing the base class
interfaces. Each class provide direct mapping to a classical step in the
optimisation workflow. \label{fig:classes}}
\end{figure}

The currently implemented subclasses for the model, problem, and cost
classes are listed in \autoref{tab:subclasses}. The cost functions are
grouped by problem type, while the model and optimiser classes can be
selected in combination with any problem-cost pair.

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.4458}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.2048}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.3494}}@{}}
\caption{List of available model, problem and cost (or likelihood)
classes. \label{tab:subclasses}}\tabularnewline
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Battery Models
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Problem Types
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Cost / Likelihood Functions
\end{minipage} \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Battery Models
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Problem Types
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Cost / Likelihood Functions
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Single particle model (SPM) & Fitting problem & Sum squared error \\
SPM with electrolyte (SPMe) & & Root mean squared error \\
Doyle-Fuller-Newman (DFN) & & Minkowski \\
Many particle model (MPM) & & Sum of power \\
Multi-species multi-reaction (MSMR) & & Gaussian log likelihood \\
Weppner Huggins & & Maximum a posteriori \\
Equivalent circuit model (ECM) & Observer & Unscented Kalman filter \\
& Design problem & Gravimetric energy density \\
& & Volumetric energy density \\
\end{longtable}

Similarly, the current algorithms available for optimisation tasks are
presented in \autoref{tab:optimisers}. From now on, the point-based
parameterisation and design optimisation tasks will simply be referred
to as optimisation tasks. This simplification can be justified by
examining \autoref{eqn:parameterisation} and \autoref{eqn:design} and
confirming that deterministic parameterisation can be viewed as an
optimisation task to minimise a distance-based cost function.

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.4062}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.3333}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.2604}}@{}}
\caption{The currently supported optimisation algorithms classified by
candidate solution type, including gradient information. (*) Scipy
minimize has gradient and non-gradient methods.
\label{tab:optimisers}}\tabularnewline
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Gradient-based
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Evolutionary
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
(Meta)heuristic
\end{minipage} \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Gradient-based
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Evolutionary
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
(Meta)heuristic
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Weight decayed adaptive moment estimation (AdamW) & Covariance matrix
adaptation (CMA-ES) & Particle swarm (PSO) \\
Improved resilient backpropagation (iRProp-) & Exponential natural
(xNES) & Nelder-Mead \\
Gradient descent & Separable natural (sNES) & Cuckoo search \\
SciPy minimize (*) & SciPy differential evolution & \\
& & \\
\end{longtable}

As discussed above, \texttt{PyBOP} provides Bayesian inference methods
such as maximum a posteriori (MAP) alongside the point-based methods in
\autoref{tab:subclasses}; however, for a full Bayesian framework, Monte
Carlo sampling is implemented within \texttt{PyBOP}. These methods
construct a posterior distribution on the inference parameters, which
can be used for uncertainty and practical identifiability. The
individual sampler classes are currently composed within \texttt{PyBOP}
from the \texttt{PINTS} library, with a base sampling class implemented
for interoperability and direct integration with the \texttt{PyBOP}
model, problem, and likelihood classes. The currently supported samplers
are listed in \autoref{tab:samplers}.

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1810}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.2155}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1379}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.2069}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.2586}}@{}}
\caption{Sampling methods supported by PyBOP, classified according to
the proposed method. \label{tab:samplers}}\tabularnewline
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Gradient-based
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Adaptive
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Slicing
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Evolutionary
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Other
\end{minipage} \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Gradient-based
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Adaptive
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Slicing
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Evolutionary
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Other
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Monomial Gamma & Delayed Rejection Adaptive & Rank Shrinking &
Differential Evolution & Metropolis Random Walk \\
No-U-Turn & Haario Bardenet & Doubling & & Emcee Hammer \\
Hamiltonian & Haario & Stepout & & Metropolis Adjusted Langevin \\
Relativistic & Rao Blackwell & & & \\
\end{longtable}

\section{Background}\label{background}

\subsection{Battery models}\label{battery-models}

In general, battery models can be written in the form of a
differential-algebraic system of equations: \begin{equation}
\frac{\mathrm{d} \mathbf{x}}{\mathrm{d} t} = f(t,\mathbf{x},\mathbf{y},\mathbf{u}(t),\mathbf{\theta}),
\label{dynamics}
\end{equation} \begin{equation}
\mathbf{y}(t) = g(t,\mathbf{x},\mathbf{y},\mathbf{u}(t),\mathbf{\theta}),
\label{output}
\end{equation} with initial conditions \begin{equation}
\mathbf{x}(0) = \mathbf{x}_0(\mathbf{\theta}).
\label{initial_conditions}
\end{equation}

Here, \(t\) is time, \(\mathbf{x}(t)\) are the (spatially discretised)
states, \(\mathbf{y}(t)\) are the outputs (e.g.~the terminal voltage),
\(\mathbf{u}(t)\) are the inputs (e.g.~the applied current) and
\(\mathbf{\theta}\) are the unknown parameters.

Common battery models include various types of equivalent circuit models
(e.g.~the Thévenin model), the Doyle--Fuller--Newman (DFN) model
(\citeproc{ref-Doyle:1993}{Doyle et al., 1993};
\citeproc{ref-Fuller:1994}{Fuller et al., 1994}) based on porous
electrode theory and its reduced-order variants including the single
particle model (SPM) (\citeproc{ref-Planella:2022}{Planella et al.,
2022}), and the multi-species, multi-reaction (MSMR) model
(\citeproc{ref-Verbrugge:2017}{Verbrugge et al., 2017}). Simplified
models that retain acceptable predictive capabilities at a lower
computational cost are widely used, for example in battery management
systems, while physics-based models are required to understand the
impact of physical parameters on battery performance. This separation of
complexity traditionally results in multiple parameterisations for a
single battery type, depending on the model structure.

\section{Examples}\label{examples}

\subsection{Parameterisation}\label{parameterisation}

The parameterisation of battery models is challenging due to the large
number of parameters that need to be identified compared to the
measurable outputs (\citeproc{ref-Andersson:2022}{Andersson et al.,
2022}; \citeproc{ref-Miguel:2021}{Miguel et al., 2021};
\citeproc{ref-Wang:2022}{Wang et al., 2022}). A complete
parameterisation often requires a stepwise identification of smaller
sets of parameters from a variety of different data sets
(\citeproc{ref-Chen:2020}{Chen et al., 2020};
\citeproc{ref-Chu:2019}{Chu et al., 2019}; \citeproc{ref-Kirk:2022}{Kirk
et al., 2023}) and excitations.

A generic data fitting optimisation problem may be formulated as:
\begin{equation}
\min_{\mathbf{\theta}} ~ \mathcal{L}_{(\mathbf{y}_i)}(\mathbf{\theta}) ~~~
\textrm{subject to equations (\ref{dynamics})\textrm{-}(\ref{initial_conditions})}
\label{eqn:parameterisation}
\end{equation}

where \(\mathcal{L} : \mathbf{\theta} \mapsto [0,\infty)\) is a cost (or
likelihood) function that quantifies the agreement between the model and
a sequence of observations \((\mathbf{y}_i)\) measured at times \(t_i\).
For gradient-based optimisers, the Jacobian of the cost function with
respect to the unknown parameters,
\((\frac{\partial \mathcal{L}}{\partial \theta})\) is computed for step
size and directional information.

Next, we demonstrate the fitting of synthetic data where the system
parameters are known. In this example problem, we use \texttt{PyBaMM}'s
implementation of the single particle model (SPM) with an added contact
resistance submodel. We assume that the battery model is already
parameterised except for two dynamic parameters, namely the lithium
diffusivity of the negative electrode active material particle (denoted
``negative particle diffusivity'') and the contact resistance. We
generate synthetic data from a one-hour discharge from 100\% state of
charge, to 0\% (denoted as 1C rate), followed by 30 minutes of
relaxation. This data is then corrupted with zero mean Gaussian noise of
amplitude 2mV, shown by the dots in
\autoref{fig:inference-time-landscape} (left). The initial states are
assumed known, although such an assumption is not generally necessary.
The underlying cost landscape explored by the optimiser is shown in
\autoref{fig:inference-time-landscape} (right).

\begin{figure}
\centering
\includegraphics[width=1\textwidth,height=\textheight]{figures/joss/sim-landscape.png}
\caption{The cost landscape for the parameterisation problem with a root
mean squared error cost function. \label{fig:inference-time-landscape}}
\end{figure}

As gradient information is available for this problem, the choice of
distance-based cost function and optimiser is not constrained. Due to
the different magnitudes of the two parameters, we apply the logarithmic
parameter transformation offered by \texttt{PyBOP}. This transforms the
optimisers search space of the optimiser to allow for a common step size
between the parameters, which is generally is not required, but improves
convergence in this problem. As a demonstration of the parameterisation
capabilities of \texttt{PyBOP}, \autoref{fig:convergence-min-max} (left)
shows the rate of convergence for each of the distance-minimising cost
functions, while \autoref{fig:convergence-min-max} (right) shows
analogous results for maximising a likelihood. The optimisation is
performed with SciPy Minimize using the gradient-based L-BFGS-B method.

\begin{figure}
\centering
\includegraphics[width=1\textwidth,height=\textheight]{figures/joss/converge.png}
\caption{Convergence in the likelihood functions obtained using various
likelihood functions and the L-BFGS-B algorithm.
\label{fig:convergence-min-max}}
\end{figure}

Next, the performance of the various optimisation algorithms is
presented by category: gradient-based in
\autoref{fig:optimiser-inference} (left), evolutionary strategies in
\autoref{fig:optimiser-inference} (middle) and (meta)heuristics in
\autoref{fig:optimiser-inference} (right) for a mean squared error cost
function. Note that the performance of the optimiser depends on the cost
environment, prior information and corresponding hyperparameters for
each specific problem.

\begin{figure}
\centering
\includegraphics[width=1\textwidth,height=\textheight]{figures/joss/optimisers_parameters.png}
\caption{Convergence in the parameter values obtained for the various
optimisation algorithms provided by \texttt{PyBOP}.
\label{fig:optimiser-inference}}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=1\textwidth,height=\textheight]{figures/joss/contour_total.png}
\caption{Cost landscape contour plot with corresponding optimisation
traces. The top row represents the gradient-based optimisers, the middle
row is the evolution-based, and the bottom row is the (meta)heuristics.
The order from left to right corresponds to the entries in
\autoref{tab:optimisers}. \label{fig:optimiser-inference}}
\end{figure}

This parameterisation task can also be approached from a Bayesian
perspective, which we will present below using PyBOP's sampler methods.
The optimisation equation presented in equation
\autoref{eqn:parameterisation} does not represent the Bayesian parameter
identification task, and as such we introduce the Bayes theorem as,

\begin{equation}
P(\theta|D) = \frac{P(D|\theta)P(\theta)}{P(D)}
\label{eqn:bayes_theorem}
\end{equation}

where, \(P(\theta|D)\) is the posterior and represents the probability
density function of the parameter. \(P(D|\theta)\) is the likelihood
function and assesses the parameter values alongside a noise model.
\(P(\theta)\) encapsulates the prior knowledge about the parameters, and
finally \(P(D)\) is the model evidence and acts as a normalising
constant so that the final posterior is a correctly scaled density
function. Our goal in parameter inference is to identify the parameter
values with the highest probability, which can be represented as a
point-based metric or as the posterior distribution, which provides
additional information about the uncertainty of the identified
parameters. Monte Carlo sampling methods are available to obtain this
posterior distribution. These methods sample from the posterior using a
variety of methods, including gradient-based methods such as No-U-Turn
(\citeproc{ref-NUTS:2011}{Hoffman \& Gelman, 2011}) and Hamiltonian
(\citeproc{ref-Hamiltonian:2011}{Brooks et al., 2011}), as well as
heuristic methods such as Differential Evolution
(\citeproc{ref-DiffEvolution:2006}{Braak, 2006}), and finally
conventional methods based on random sampling with rejection criteria
(\citeproc{ref-metropolis:1953}{Metropolis et al., 1953}). PyBOP offers
a sampling class that provides an interface to these samplers, which are
supported by the Probabilistic Inference of Noise Time-Series (PINTS)
package. \autoref{fig:posteriors} below shows the sampled posterior for
the synthetic workflow described above, using an adaptive
covariance-based sampler, Haario Bardenet
(\citeproc{ref-Haario:2001}{Haario et al., 2001}).

\begin{figure}
\centering
\includegraphics[width=1\textwidth,height=\textheight]{figures/joss/posteriors.png}
\caption{Posterior distributions for model parameters alongside
identified noise on the observations. Shaded area denotes confidence
bounds for each parameter. \label{fig:posteriors}}
\end{figure}

\subsection{Design optimisation}\label{design-optimisation}

Design optimisation is supported within \texttt{PyBOP} to guide future
battery design development by identifying parameter sensitivities that
can unlock improvements in battery performance. This problem can be
viewed in a similar way to the parameterisation workflows described
above, but with the aim of maximising a distance metric rather than
minimising it. In the case of design optimisation to maximise
gravimetric energy density, \texttt{PyBOP} minimises the negative of the
cost function, where the cost metric is no longer a distance between two
time series vectors, but the integrated energy of the vector normalised
by with the corresponding cell mass. This is typically quantified for
operating conditions such as a 1C discharge, at a given temperature.

Design optimisation can be written in the form of a constrained
optimisation problem as: \begin{equation}
\min_{\mathbf{\theta} \in \Omega} ~ \mathcal{L}(\mathbf{\theta}) ~~~
\textrm{subject to equations (\ref{dynamics})\textrm{-}(\ref{initial_conditions})}
\label{eqn:design}
\end{equation}

where \(\mathcal{L} : \mathbf{\theta} \mapsto [0,\infty)\) is a cost
function that quantifies the desirability of the design and \(\Omega\)
is the set of allowable parameter values.

As an example, we consider the problem of maximising the gravimetric
energy density subject to constraints on two of the geometric electrode
parameters (\citeproc{ref-Couto:2023}{Couto et al., 2023}). For this
example, we use the\texttt{PyBaMM} implementation of the single particle
model with electrolyte (SPMe) to investigate the effect of the positive
electrode thickness and the active material volume fraction on the
target cost. Since the active material volume fraction is related to the
electrode porosity, the porosity is defined with a driven constraint
from the volume fraction. In this problem, we estimate the 1C rate from
the theoretical capacity for each iteration of the design. For this
example, we employ the Particle Swarm Optimisation (PSO) algorithm.

\begin{figure}
\centering
\includegraphics[width=1\textwidth,height=\textheight]{figures/joss/design.png}
\caption{The gravimetric landscape alongside the corresponding initial
and optimised voltage profiles for a 1C discharge.
\label{fig:design_gravimetric}}
\end{figure}

\autoref{fig:design_gravimetric} (left) shows the optimiser's search
over the gravimetric energy density parameter space. The predicted
improvement in the discharge profile between the initial and optimised
parameter values (right) for their respective applied 1C current.

\section{Acknowledgements}\label{acknowledgements}

We gratefully acknowledge all
\href{https://github.com/pybop-team/PyBOP?tab=readme-ov-file\#contributors-}{contributors}
to this package. This work was supported by the Faraday Institution
Multiscale Modelling (MSM) project (ref. FIRG059), UKRI's Horizon Europe
Guarantee (ref. 10038031), and EU IntelLiGent project (ref. 101069765).

\section*{References}\label{references}
\addcontentsline{toc}{section}{References}

\phantomsection\label{refs}
\begin{CSLReferences}{1}{0}
\bibitem[\citeproctext]{ref-Andersson:2022}
Andersson, M., Streb, M., Ko, J. Y., Löfqvist Klass, V., Klett, M.,
Ekström, H., Johansson, M., \& Lindbergh, G. (2022). Parametrization of
physics-based battery models from input-output data: {A} review of
methodology and current research. \emph{Journal of Power Sources},
\emph{521}(November 2021), 230859.
\url{https://doi.org/10.1016/j.jpowsour.2021.230859}

\bibitem[\citeproctext]{ref-DiffEvolution:2006}
Braak, C. J. F. T. (2006). A {Markov} {Chain} {Monte} {Carlo} version of
the genetic algorithm {Differential} {Evolution}: Easy {Bayesian}
computing for real parameter spaces. \emph{Statistics and Computing},
\emph{16}(3), 239--249. \url{https://doi.org/10.1007/s11222-006-8769-1}

\bibitem[\citeproctext]{ref-jax:2018}
Bradbury, J., Frostig, R., Hawkins, P., Johnson, M. J., Leary, C.,
Maclaurin, D., Necula, G., Paszke, A., VanderPlas, J., Wanderman-Milne,
S., \& Zhang, Q. (2018). \emph{{JAX}: Composable transformations of
{P}ython+{N}um{P}y programs} (Version 0.3.13).
\url{http://github.com/jax-ml/jax}

\bibitem[\citeproctext]{ref-Hamiltonian:2011}
Brooks, S., Gelman, A., Jones, G., \& Meng, X.-L. (2011). \emph{Handbook
of markov chain monte carlo}. Chapman; Hall/CRC.
\url{https://doi.org/10.1201/b10905}

\bibitem[\citeproctext]{ref-blackjax:2024}
Cabezas, A., Corenflos, A., Lao, J., \& Louf, R. (2024). \emph{BlackJAX:
Composable {B}ayesian inference in {JAX}}.
\url{https://arxiv.org/abs/2402.10797}

\bibitem[\citeproctext]{ref-Chen:2020}
Chen, C.-H., Brosa Planella, F., O'Regan, K., Gastol, D., Widanage, W.
D., \& Kendrick, E. (2020). Development of experimental techniques for
parameterization of multi-scale lithium-ion battery models.
\emph{Journal of The Electrochemical Society}, \emph{167}(8), 080534.
\url{https://doi.org/10.1149/1945-7111/ab9050}

\bibitem[\citeproctext]{ref-Chu:2019}
Chu, Z., Plett, G. L., Trimboli, M. S., \& Ouyang, M. (2019). A
control-oriented electrochemical model for lithium-ion battery, {P}art
{I}: {L}umped-parameter reduced-order model with constant phase element.
\emph{Journal of Energy Storage}, \emph{25}(August), 100828.
\url{https://doi.org/10.1016/j.est.2019.100828}

\bibitem[\citeproctext]{ref-Clerx:2019}
Clerx, M., Robinson, M., Lambert, B., Lei, C. L., Ghosh, S., Mirams, G.
R., \& Gavaghan, D. J. (2019). Probabilistic inference on noisy time
series ({PINTS}). \emph{Journal of Open Research Software}, \emph{7}(1),
23. \url{https://doi.org/10.5334/jors.252}

\bibitem[\citeproctext]{ref-Couto:2023}
Couto, L. D., Charkhgard, M., Karaman, B., Job, N., \& Kinnaert, M.
(2023). Lithium-ion battery design optimization based on a dimensionless
reduced-order electrochemical model. \emph{Energy}, \emph{263}(PE),
125966. \url{https://doi.org/10.1016/j.energy.2022.125966}

\bibitem[\citeproctext]{ref-optax:2020}
DeepMind, Babuschkin, I., Baumli, K., Bell, A., Bhupatiraju, S., Bruce,
J., Buchlovsky, P., Budden, D., Cai, T., Clark, A., Danihelka, I.,
Dedieu, A., Fantacci, C., Godwin, J., Jones, C., Hemsley, R., Hennigan,
T., Hessel, M., Hou, S., \ldots{} Viola, F. (2020). \emph{The
{D}eep{M}ind {JAX} {E}cosystem}. \url{http://github.com/google-deepmind}

\bibitem[\citeproctext]{ref-Doyle:1993}
Doyle, M., Fuller, T. F., \& Newman, J. (1993). {Modeling of
Galvanostatic Charge and Discharge of the Lithium/Polymer/Insertion
Cell}. \emph{Journal of The Electrochemical Society}, \emph{140}(6),
1526--1533. \url{https://doi.org/10.1149/1.2221597}

\bibitem[\citeproctext]{ref-Fuller:1994}
Fuller, T. F., Doyle, M., \& Newman, J. (1994). Simulation and
optimization of the dual lithium ion insertion cell. \emph{Journal of
The Electrochemical Society}, \emph{141}(1), 1.
\url{https://doi.org/10.1149/1.2054684}

\bibitem[\citeproctext]{ref-Haario:2001}
Haario, H., Saksman, E., \& Tamminen, J. (2001). An {Adaptive}
{Metropolis} {Algorithm}. \emph{Bernoulli}, \emph{7}(2), 223.
\url{https://doi.org/10.2307/3318737}

\bibitem[\citeproctext]{ref-NUTS:2011}
Hoffman, M. D., \& Gelman, A. (2011). \emph{The no-u-turn sampler:
Adaptively setting path lengths in hamiltonian monte carlo}.
\url{https://arxiv.org/abs/1111.4246}

\bibitem[\citeproctext]{ref-Kirk:2022}
Kirk, T. L., Lewis-Douglas, A., Howey, D., Please, C. P., \& Jon
Chapman, S. (2023). Nonlinear electrochemical impedance spectroscopy for
lithium-ion battery model parameterization. \emph{Journal of The
Electrochemical Society}, \emph{170}(1), 010514.
\url{https://doi.org/10.1149/1945-7111/acada7}

\bibitem[\citeproctext]{ref-BPX:2023}
Korotkin, I., Timms, R., Foster, J. F., Dickinson, E., \& Robinson, M.
(2023). Battery parameter eXchange. In \emph{GitHub repository}. The
Faraday Institution. \url{https://github.com/FaradayInstitution/BPX}

\bibitem[\citeproctext]{ref-metropolis:1953}
Metropolis, N., Rosenbluth, A. W., Rosenbluth, M. N., Teller, A. H., \&
Teller, E. (1953). Equation of {State} {Calculations} by {Fast}
{Computing} {Machines}. \emph{The Journal of Chemical Physics},
\emph{21}(6), 1087--1092. \url{https://doi.org/10.1063/1.1699114}

\bibitem[\citeproctext]{ref-Miguel:2021}
Miguel, E., Plett, G. L., Trimboli, M. S., Oca, L., Iraola, U., \&
Bekaert, E. (2021). Review of computational parameter estimation methods
for electrochemical models. \emph{Journal of Energy Storage},
\emph{44}(PB), 103388. \url{https://doi.org/10.1016/j.est.2021.103388}

\bibitem[\citeproctext]{ref-numpyro:2019}
Phan, D., Pradhan, N., \& Jankowiak, M. (2019). Composable effects for
flexible and accelerated probabilistic programming in NumPyro.
\emph{arXiv Preprint arXiv:1912.11554}.

\bibitem[\citeproctext]{ref-Planella:2022}
Planella, F. B., Ai, W., Boyce, A. M., Ghosh, A., Korotkin, I., Sahu,
S., Sulzer, V., Timms, R., Tranter, T. G., Zyskin, M., Cooper, S. J.,
Edge, J. S., Foster, J. M., Marinescu, M., Wu, B., \& Richardson, G.
(2022). {A Continuum of Physics-Based Lithium-Ion Battery Models
Reviewed}. \emph{Progress in Energy}, \emph{4}(4), 042003.
\url{https://doi.org/10.1088/2516-1083/ac7d31}

\bibitem[\citeproctext]{ref-Sulzer:2021}
Sulzer, V., Marquis, S. G., Timms, R., Robinson, M., \& Chapman, S. J.
(2021). {Python Battery Mathematical Modelling (PyBaMM)}. \emph{Journal
of Open Research Software}, \emph{9}(1), 14.
\url{https://doi.org/10.5334/jors.309}

\bibitem[\citeproctext]{ref-Tranter2022}
Tranter, T. G., Timms, R., Sulzer, V., Planella, F. B., Wiggins, G. M.,
Karra, S. V., Agarwal, P., Chopra, S., Allu, S., Shearing, P. R., \&
Brett, D. J. l. (2022). Liionpack: A python package for simulating packs
of batteries with PyBaMM. \emph{Journal of Open Source Software},
\emph{7}(70), 4051. \url{https://doi.org/10.21105/joss.04051}

\bibitem[\citeproctext]{ref-Verbrugge:2017}
Verbrugge, M., Baker, D., Koch, B., Xiao, X., \& Gu, W. (2017).
Thermodynamic model for substitutional materials: Application to
lithiated graphite, spinel manganese oxide, iron phosphate, and layered
nickel-manganese-cobalt oxide. \emph{Journal of The Electrochemical
Society}, \emph{164}(11), E3243.
\url{https://doi.org/10.1149/2.0341708jes}

\bibitem[\citeproctext]{ref-SciPy:2020}
Virtanen, P., Gommers, R., Oliphant, T. E., Haberland, M., Reddy, T.,
Cournapeau, D., Burovski, E., Peterson, P., Weckesser, W., Bright, J.,
van der Walt, S. J., Brett, M., Wilson, J., Millman, K. J., Mayorov, N.,
Nelson, A. R. J., Jones, E., Kern, R., Larson, E., \ldots{} SciPy 1.0
Contributors. (2020). {{SciPy} 1.0: Fundamental Algorithms for
Scientific Computing in Python}. \emph{Nature Methods}, \emph{17},
261--272. \url{https://doi.org/10.1038/s41592-019-0686-2}

\bibitem[\citeproctext]{ref-Wang:2022}
Wang, A. A., O'Kane, S. E. J., Brosa Planella, F., Houx, J. L., O'Regan,
K., Zyskin, M., Edge, J., Monroe, C. W., Cooper, S. J., Howey, D. A.,
Kendrick, E., \& Foster, J. M. (2022). Review of parameterisation and a
novel database {(LiionDB)} for continuum {Li-ion} battery models.
\emph{Progress in Energy}, \emph{4}(3), 032004.
\url{https://doi.org/10.1088/2516-1083/ac692c}

\end{CSLReferences}

\end{document}