### 2024-12-09 joss/paper.preprint.tex
feat: updates figures, contour plot into plotly subplots.

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% definitions for citeproc citations
\NewDocumentCommand\citeproctext{}{}
\NewDocumentCommand\citeproc{mm}{%
  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
\makeatletter
 % allow citations to break across lines
 \let\@cite@ofmt\@firstofone
 % avoid brackets around text for \cite:
 \def\@biblabel#1{}
 \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
 {\begin{list}{}{%
  \setlength{\itemindent}{0pt}
  \setlength{\leftmargin}{0pt}
  \setlength{\parsep}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
   \setlength{\leftmargin}{\cslhangindent}
   \setlength{\itemindent}{-1\cslhangindent}
  \fi
  % set entry spacing
  \setlength{\itemsep}{#2\baselineskip}}}
 {\end{list}}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\ifLuaTeX
\usepackage[bidi=basic]{babel}
\else
\usepackage[bidi=default]{babel}
\fi
\babelprovide[main,import]{american}
% get rid of language-specific shorthands (see #6817):
\let\LanguageShortHands\languageshorthands
\def\languageshorthands#1{}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={PyBOP: A Python package for battery model optimisation and
parameterisation},
  pdfauthor={Brady Planden, Nicola E. Courtier, Martin Robinson, Agriya
Khetarpal, Ferran Brosa Planella, David A. Howey},
  pdflang={en-US},
  colorlinks=true,
  linkcolor={Maroon},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{PyBOP: A Python package for battery model optimisation and
parameterisation}

\definecolor{c53baa1}{RGB}{83,186,161}
\definecolor{c202826}{RGB}{32,40,38}
\def \rorglobalscale {0.1}
\newcommand{\rorlogo}{%
\begin{tikzpicture}[y=1cm, x=1cm, yscale=\rorglobalscale,xscale=\rorglobalscale, every node/.append style={scale=\rorglobalscale}, inner sep=0pt, outer sep=0pt]
  \begin{scope}[even odd rule,line join=round,miter limit=2.0,shift={(0,3.0889892500000005)},xscale=0.9943,yscale=0.9894]
    \path[fill=white,even odd rule,line join=round,miter limit=2.0] (0.0, 3.1221) rectangle (4.364, -0.0336);
  \end{scope}
  \begin{scope}[even odd rule,line join=round,miter limit=2.0,shift={(-0.025, 0.0216)}]
    \path[fill=c53baa1,nonzero rule,line join=round,miter limit=2.0] (1.8164, 3.012) -- (1.4954, 2.5204) -- (1.1742, 3.012) -- (1.8164, 3.012) -- cycle;
    \path[fill=c53baa1,nonzero rule,line join=round,miter limit=2.0] (3.1594, 3.012) -- (2.8385, 2.5204) -- (2.5172, 3.012) -- (3.1594, 3.012) -- cycle;
    \path[fill=c53baa1,nonzero rule,line join=round,miter limit=2.0] (1.1742, 0.0669) -- (1.4954, 0.5588) -- (1.8164, 0.0669) -- (1.1742, 0.0669) -- cycle;
    \path[fill=c53baa1,nonzero rule,line join=round,miter limit=2.0] (2.5172, 0.0669) -- (2.8385, 0.5588) -- (3.1594, 0.0669) -- (2.5172, 0.0669) -- cycle;
    \path[fill=c202826,nonzero rule,line join=round,miter limit=2.0] (3.8505, 1.4364).. controls (3.9643, 1.4576) and (4.0508, 1.5081) .. (4.1098, 1.5878).. controls (4.169, 1.6674) and (4.1984, 1.7642) .. (4.1984, 1.8777).. controls (4.1984, 1.9719) and (4.182, 2.0503) .. (4.1495, 2.1132).. controls (4.1169, 2.1762) and (4.0727, 2.2262) .. (4.0174, 2.2635).. controls (3.9621, 2.3006) and (3.8976, 2.3273) .. (3.824, 2.3432).. controls (3.7505, 2.359) and (3.6727, 2.367) .. (3.5909, 2.367) -- (2.9676, 2.367) -- (2.9676, 1.8688).. controls (2.9625, 1.8833) and (2.9572, 1.8976) .. (2.9514, 1.9119).. controls (2.9083, 2.0164) and (2.848, 2.1056) .. (2.7705, 2.1791).. controls (2.6929, 2.2527) and (2.6014, 2.3093) .. (2.495, 2.3487).. controls (2.3889, 2.3881) and (2.2728, 2.408) .. (2.1468, 2.408).. controls (2.0209, 2.408) and (1.905, 2.3881) .. (1.7986, 2.3487).. controls (1.6925, 2.3093) and (1.6007, 2.2527) .. (1.5232, 2.1791).. controls (1.4539, 2.1132) and (1.3983, 2.0346) .. (1.3565, 1.9436).. controls (1.3504, 2.009) and (1.3351, 2.0656) .. (1.3105, 2.1132).. controls (1.2779, 2.1762) and (1.2338, 2.2262) .. (1.1785, 2.2635).. controls (1.1232, 2.3006) and (1.0586, 2.3273) .. (0.985, 2.3432).. controls (0.9115, 2.359) and (0.8337, 2.367) .. (0.7519, 2.367) -- (0.1289, 2.367) -- (0.1289, 0.7562) -- (0.4837, 0.7562) -- (0.4837, 1.4002) -- (0.6588, 1.4002) -- (0.9956, 0.7562) -- (1.4211, 0.7562) -- (1.0118, 1.4364).. controls (1.1255, 1.4576) and (1.2121, 1.5081) .. (1.2711, 1.5878).. controls (1.2737, 1.5915) and (1.2761, 1.5954) .. (1.2787, 1.5991).. controls (1.2782, 1.5867) and (1.2779, 1.5743) .. (1.2779, 1.5616).. controls (1.2779, 1.4327) and (1.2996, 1.3158) .. (1.3428, 1.2113).. controls (1.3859, 1.1068) and (1.4462, 1.0176) .. (1.5237, 0.944).. controls (1.601, 0.8705) and (1.6928, 0.8139) .. (1.7992, 0.7744).. controls (1.9053, 0.735) and (2.0214, 0.7152) .. (2.1474, 0.7152).. controls (2.2733, 0.7152) and (2.3892, 0.735) .. (2.4956, 0.7744).. controls (2.6016, 0.8139) and (2.6935, 0.8705) .. (2.771, 0.944).. controls (2.8482, 1.0176) and (2.9086, 1.1068) .. (2.952, 1.2113).. controls (2.9578, 1.2253) and (2.9631, 1.2398) .. (2.9681, 1.2544) -- (2.9681, 0.7562) -- (3.3229, 0.7562) -- (3.3229, 1.4002) -- (3.4981, 1.4002) -- (3.8349, 0.7562) -- (4.2603, 0.7562) -- (3.8505, 1.4364) -- cycle(0.9628, 1.7777).. controls (0.9438, 1.7534) and (0.92, 1.7357) .. (0.8911, 1.7243).. controls (0.8623, 1.7129) and (0.83, 1.706) .. (0.7945, 1.7039).. controls (0.7588, 1.7015) and (0.7252, 1.7005) .. (0.6932, 1.7005) -- (0.4839, 1.7005) -- (0.4839, 2.0667) -- (0.716, 2.0667).. controls (0.7477, 2.0667) and (0.7805, 2.0643) .. (0.8139, 2.0598).. controls (0.8472, 2.0553) and (0.8768, 2.0466) .. (0.9025, 2.0336).. controls (0.9282, 2.0206) and (0.9496, 2.0021) .. (0.9663, 1.9778).. controls (0.9829, 1.9534) and (0.9914, 1.9209) .. (0.9914, 1.8799).. controls (0.9914, 1.8362) and (0.9819, 1.8021) .. (0.9628, 1.7777) -- cycle(2.6125, 1.3533).. controls (2.5889, 1.2904) and (2.5553, 1.2359) .. (2.5112, 1.1896).. controls (2.4672, 1.1433) and (2.4146, 1.1073) .. (2.3529, 1.0814).. controls (2.2916, 1.0554) and (2.2228, 1.0427) .. (2.1471, 1.0427).. controls (2.0712, 1.0427) and (2.0026, 1.0557) .. (1.9412, 1.0814).. controls (1.8799, 1.107) and (1.8272, 1.1433) .. (1.783, 1.1896).. controls (1.7391, 1.2359) and (1.7052, 1.2904) .. (1.6817, 1.3533).. controls (1.6581, 1.4163) and (1.6465, 1.4856) .. (1.6465, 1.5616).. controls (1.6465, 1.6359) and (1.6581, 1.705) .. (1.6817, 1.7687).. controls (1.7052, 1.8325) and (1.7388, 1.8873) .. (1.783, 1.9336).. controls (1.8269, 1.9799) and (1.8796, 2.0159) .. (1.9412, 2.0418).. controls (2.0026, 2.0675) and (2.0712, 2.0804) .. (2.1471, 2.0804).. controls (2.223, 2.0804) and (2.2916, 2.0675) .. (2.3529, 2.0418).. controls (2.4143, 2.0161) and (2.467, 1.9799) .. (2.5112, 1.9336).. controls (2.5551, 1.8873) and (2.5889, 1.8322) .. (2.6125, 1.7687).. controls (2.636, 1.705) and (2.6477, 1.6359) .. (2.6477, 1.5616).. controls (2.6477, 1.4856) and (2.636, 1.4163) .. (2.6125, 1.3533) -- cycle(3.8015, 1.7777).. controls (3.7825, 1.7534) and (3.7587, 1.7357) .. (3.7298, 1.7243).. controls (3.701, 1.7129) and (3.6687, 1.706) .. (3.6333, 1.7039).. controls (3.5975, 1.7015) and (3.5639, 1.7005) .. (3.5319, 1.7005) -- (3.3226, 1.7005) -- (3.3226, 2.0667) -- (3.5547, 2.0667).. controls (3.5864, 2.0667) and (3.6192, 2.0643) .. (3.6526, 2.0598).. controls (3.6859, 2.0553) and (3.7155, 2.0466) .. (3.7412, 2.0336).. controls (3.7669, 2.0206) and (3.7883, 2.0021) .. (3.805, 1.9778).. controls (3.8216, 1.9534) and (3.8301, 1.9209) .. (3.8301, 1.8799).. controls (3.8301, 1.8362) and (3.8206, 1.8021) .. (3.8015, 1.7777) -- cycle;
  \end{scope}
\end{tikzpicture}
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Authors and Affiliations

\usepackage[affil-it]{authblk}
\usepackage{orcidlink}
%% \renewcommand\Authsep{, }
\setlength{\affilsep}{1em}
\author[1%
  %
  \ensuremath\mathparagraph]{Brady Planden%
    \,\orcidlink{0000-0002-1082-9125}\,%
    }
\author[1,2%
  %
  ]{Nicola E. Courtier%
    \,\orcidlink{0000-0002-5714-1096}\,%
    }
\author[3%
  %
  ]{Martin Robinson%
    \,\orcidlink{0000-0002-1572-6782}\,%
    }
\author[4%
  %
  ]{Agriya Khetarpal%
    \,\orcidlink{0000-0002-1112-1786}\,%
    }
\author[2,5%
  %
  ]{Ferran Brosa Planella%
    \,\orcidlink{0000-0001-6363-2812}\,%
    }
\author[1,2%
  %
  ]{David A. Howey%
    \,\orcidlink{0000-0002-0620-3955}\,%
    }

\affil[1]{Department of Engineering Science, University of Oxford,
Oxford, UK%
  }
\affil[2]{The Faraday Institution, Harwell Campus, Didcot, UK%
  }
\affil[3]{Research Software Engineering Group, University of Oxford,
Oxford, UK%
  }
\affil[4]{Quansight Labs%
  }
\affil[5]{Mathematics Institute, University of Warwick, Coventry, UK%
  }
\affil[$\mathparagraph$]{Corresponding author: %
   %
}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\date{04 December 2024}

\begin{document}
\maketitle

\section{Summary}\label{summary}

The Python Battery Optimisation and Parameterisation (\texttt{PyBOP})
package provides methods for estimating and optimising battery model
parameters, offering both deterministic and stochastic approaches with
example workflows to assist users.

\texttt{PyBOP} enables parameter identification from data for various
battery models, including both the electrochemical and equivalent
circuit models provided by the popular open-source \texttt{PyBaMM}
package (\citeproc{ref-Sulzer:2021}{Sulzer et al., 2021}). Using the
same workflow, \texttt{PyBOP} can also be used for design optimisation
under user-defined operating conditions across a variety of model
structures and design goals. \texttt{PyBOP} facilitates optimisation
using a range of methods, with diagnostics for examining optimiser
performance and convergence of the cost and corresponding parameters.
Identified parameters can be used for prediction, on-line control, and
design optimisation to accelerate research and development and improve
battery utilisation.

\section{Statement of need}\label{statement-of-need}

\texttt{PyBOP} is a Python package that provides a user-friendly,
object-oriented interface for optimising battery model parameters.
\texttt{PyBOP} leverages the open-source \texttt{PyBaMM} package
(\citeproc{ref-Sulzer:2021}{Sulzer et al., 2021}) to formulate and solve
battery models. Together, these packages serve a broad audience
including students, engineers, and researchers in academia and industry,
enabling the use of advanced models where previously this was not
possible without specialised knowledge of battery modelling, parameter
inference and software development. \texttt{PyBOP} emphasises clear and
informative diagnostics and workflows to support users with varying
levels of domain expertise, and provides access to a wide range of
optimisation and sampling algorithms. These methods are provided through
interfaces to \texttt{PINTS} (\citeproc{ref-Clerx:2019}{Clerx et al.,
2019}), \texttt{SciPy} (\citeproc{ref-SciPy:2020}{Virtanen et al.,
2020}), and \texttt{PyBOP}'s own implementations of algorithms such as
adaptive moment estimation with weight decay (AdamW), gradient descent,
and Cuckoo search.

\texttt{PyBOP} supports the battery parameter exchange (BPX) standard
(\citeproc{ref-BPX:2023}{Korotkin et al., 2023}) for sharing battery
parameter sets. These parameter sets are typically costly to obtain due
to the specialised equipment and time required for characterisation
experiments, the need for battery domain knowledge, and the
computational cost of parameter estimation. \texttt{PyBOP} reduces the
requirements for the latter two categories by providing fast parameter
estimation, standardised workflows, and parameter set interoperability
(via BPX).

This package complements other lithium-ion battery modelling packages
built around \texttt{PyBaMM}, such as \texttt{liionpack} for battery
pack simulation (\citeproc{ref-Tranter2022}{Tranter et al., 2022}) and
\texttt{pybamm-eis} for fast numerical computation of the
electrochemical impedance for any battery model, since the identified
parameters from \texttt{PyBOP} are easily exportable to these other
packages.

\section{Architecture}\label{architecture}

\texttt{PyBOP} has a layered structure designed to package up the
necessary functionality to compute forward predictions, process the
results, and run the optimisation and sampling algorithms. The forward
model is solved using the popular battery modelling package,
\texttt{PyBaMM}, with construction, parameterisation, and discretisation
managed by \texttt{PyBOP}'s model interface to \texttt{PyBaMM}. This
provides a robust object construction process with a consistent
interface between models and optimisers. The statistical methods and
optimisation algorithms are constructed to interface cleanly with the
forward model predictions. Furthermore, identifiability metrics are
provided with the estimated parameters (through Hessian approximation of
the cost functions in frequentist workflows, and posterior distributions
in Bayesian workflows).

\begin{figure}
\centering
\includegraphics[width=0.8\textwidth,height=\textheight]{figures/PyBOP-high-level.pdf}
\caption{PyBOP's interface to supporting funding agencies, alongside a
visualisation of the general workflow for parameterisation and
optimisation. \label{fig:high-level}}
\end{figure}

\texttt{PyBOP} formulates the inference process into four key classes,
namely, the model, the problem, the cost/likelihood, and the
optimiser/sampler, as shown in \autoref{fig:classes}. Each of these
objects represents a base class with child classes constructing
specialised functionality for different workflows. The model class
constructs a \texttt{PyBaMM} forward model for a given set of model
equations, initial conditions, spatial discretisation, and numerical
solver. By composing \texttt{PyBaMM} directly into \texttt{PyBOP},
specialised models can be constructed alongside the standard models,
which can be modified and purposefully constructed for different
inference tasks. One such example is spatial re-discretisation, which is
required when one or more geometric parameters are being optimised. In
this situation, \texttt{PyBOP} rebuilds the \texttt{PyBaMM} model a
minimum number of times while maintaining the problem, cost, and
optimiser objects, providing improved performance benefits to users.
Alongside construction of the forward model, \texttt{PyBOP}'s model
class provides methods for obtaining sensitivities from the prediction,
enabling gradient-based optimisation algorithms. A forward prediction,
along with its corresponding sensitivities, is provided to the problem
class for processing and exception control. A standardised data
structure is then provided to the cost classes, which then computes a
distance, design, or likelihood-based metric for optimisation. For
deterministic optimisation, the optimisers minimise the cost function or
the negative log-likelihood if a likelihood class is provided. Bayesian
inference is provided by sampler classes, which accept the LogPosterior
class and sample from it using \texttt{PINTS} based Monte Carlo
algorithms at the time of submission. In the typical workflow, the
classes in \autoref{fig:classes} are constructed in sequence.

In addition to the core architecture, \texttt{PyBOP} provides several
specialised inference and optimisation processes. One such instance is
parameter inference from electrochemical impedance spectroscopy (EIS)
simulations. PyBOP discretises the EIS forward model into a sparse mass
matrix form with accompanying auto-differentiated Jacobian. These
objects are then translated into the frequency domain with a linear
solution used to compute the battery model impedance. In this situation,
the forward models are constructed within the spatial re-discretisation
workflow, allowing for geometric parameter inference from EIS
simulations. Furthermore, \texttt{PyBOP} builds on the \texttt{JAX}
(\citeproc{ref-jax:2018}{Bradbury et al., 2018}) numerical solvers
provided by \texttt{PyBaMM} by providing \texttt{JAX}-based cost
functions for automatic forward model differentiation with respect to
the parameters. This functionality provides a performance improvement,
and allows users to use many other JAX-based inference packages in order
to optimise their cost function, such as \texttt{Numpyro}
(\citeproc{ref-numpyro:2019}{Phan et al., 2019}), \texttt{BlackJAX}
(\citeproc{ref-blackjax:2024}{Cabezas et al., 2024}), and \texttt{Optax}
(\citeproc{ref-optax:2020}{DeepMind et al., 2020}).

\begin{figure}
\centering
\includegraphics[width=0.8\textwidth,height=\textheight]{figures/PyBOP_components.drawio.png}
\caption{The core \texttt{PyBOP} architecture, showcasing the base class
interfaces. Each class provides a direct mapping to a classical step in
the optimisation workflow. \label{fig:classes}}
\end{figure}

The currently implemented subclasses for the model, problem, and cost
classes are listed in \autoref{tab:subclasses}. The model and optimiser
classes can be selected in combination with any problem-cost pair.

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.4458}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.2048}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.3494}}@{}}
\caption{List of available model, problem and cost (or likelihood)
classes. \label{tab:subclasses}}\tabularnewline
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Battery Models
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Problem Types
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Cost / Likelihood Functions
\end{minipage} \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Battery Models
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Problem Types
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Cost / Likelihood Functions
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Single particle model (SPM) & Fitting problem & Sum squared error \\
SPM with electrolyte (SPMe) & Design problem & Root mean squared
error \\
Doyle-Fuller-Newman (DFN) & Observer & Minkowski \\
Many particle model (MPM) & & Sum of power \\
Multi-species multi-reaction (MSMR) & & Gaussian log likelihood \\
Weppner Huggins & & Maximum a posteriori \\
Equivalent circuit model (ECM) & & Volumetric energy density \\
& & Gravimetric energy density \\
& & Unscented Kalman filter \\
\end{longtable}

Similarly, the current algorithms available for optimisation tasks are
presented in \autoref{tab:optimisers}. It should be noted that
\texttt{SciPy} minimize comprises a range of gradient and non-gradient
methods. From now on, the point-based parameterisation and design
optimisation tasks will simply be referred to as optimisation tasks.
This simplification can be justified by examining
\autoref{eqn:parameterisation} and \autoref{eqn:design} and confirming
that deterministic parameterisation can be viewed as an optimisation
task to minimise a distance-based cost function.

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.4359}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.3333}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.2308}}@{}}
\caption{The currently supported optimisation algorithms classified by
candidate solution type, including gradient information.
\label{tab:optimisers}}\tabularnewline
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Gradient-based
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Evolutionary
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
(Meta)heuristic
\end{minipage} \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Gradient-based
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Evolutionary
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
(Meta)heuristic
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Weight decayed adaptive moment estimation (AdamW) & Covariance matrix
adaptation (CMA-ES) & Particle swarm (PSO) \\
Improved resilient backpropagation (iRProp-) & Exponential natural
(xNES) & Nelder-Mead \\
Gradient descent & Separable natural (sNES) & Cuckoo search \\
SciPy minimize & SciPy differential evolution & \\
\end{longtable}

As discussed above, in addition to point-based optimisation estimates
\autoref{tab:subclasses}, \texttt{PyBOP} also provides Monte Carlo
sampling routines in order to recover distributions of likely parameter
values. The Monte Carlo methods construct a posterior distribution on
the inference parameters, which can be used to assess uncertainty and
practical identifiability. The individual sampler classes are currently
composed within \texttt{PyBOP} from the \texttt{PINTS} library, with a
base sampler class implemented for interoperability and direct
integration with the \texttt{PyBOP} model, problem, and likelihood
classes. The currently supported samplers are listed in
\autoref{tab:samplers}.

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1624}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.2393}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1368}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.2051}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.2564}}@{}}
\caption{Sampling methods supported by \texttt{PyBOP}, classified
according to the proposed method. \label{tab:samplers}}\tabularnewline
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Gradient-based
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Adaptive
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Slicing
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Evolutionary
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Other
\end{minipage} \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Gradient-based
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Adaptive
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Slicing
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Evolutionary
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Other
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Monomial Gamma & Delayed Rejection Adaptive & Rank---Shrinking &
Differential Evolution & Metropolis Random Walk \\
No-U-Turn & Haario Bardenet & Doubling & & Emcee Hammer \\
Hamiltonian & Haario & Stepout & & Metropolis Adjusted Langevin \\
Relativistic & Rao Blackwell & & & \\
\end{longtable}

\section{Background}\label{background}

\subsection{Battery models}\label{battery-models}

In general, battery models (after spatial discretisation) can be written
in the form of a differential-algebraic system of equations:
\begin{equation}
\frac{\mathrm{d} \mathbf{x}}{\mathrm{d} t} = f(t,\mathbf{x},\mathbf{\theta}),
\label{dynamics}
\end{equation} \begin{equation}
0 = g(t, \mathbf{x}, \mathbf{\theta})
\label{algebraic}
\end{equation} \begin{equation}
\mathbf{y}(t) = h(t, \mathbf{x}, \mathbf{\theta}),
\label{output}
\end{equation} with initial conditions \begin{equation}
\mathbf{x}(0) = \mathbf{x}_0(\mathbf{\theta}).
\label{initial_conditions}
\end{equation}

Here, \(t\) is time, \(\mathbf{x}(t)\) are the (spatially discretised)
states, \(\mathbf{y}(t)\) are the outputs (e.g.~the terminal voltage)
and \(\mathbf{\theta}\) are the unknown parameters.

Common battery models include various types of equivalent circuit models
(e.g.~the Thévenin model), the Doyle--Fuller--Newman (DFN) model
(\citeproc{ref-Doyle:1993}{Doyle et al., 1993};
\citeproc{ref-Fuller:1994}{Fuller et al., 1994}) based on porous
electrode theory and its reduced-order variants including the single
particle model (SPM) (\citeproc{ref-Planella:2022}{Brosa Planella et
al., 2022}), and the multi-species, multi-reaction (MSMR) model
(\citeproc{ref-Verbrugge:2017}{Verbrugge et al., 2017}). Simplified
models that retain acceptable predictive capabilities at a lower
computational cost are widely used, for example in battery management
systems, while physics-based models are required to understand the
impact of physical parameters on battery performance. This separation of
complexity traditionally results in multiple parameterisations for a
single battery type, depending on the model structure.

\section{Examples}\label{examples}

\subsection{Parameterisation}\label{parameterisation}

The parameterisation of battery models is challenging due to the large
number of parameters that need to be identified compared to the
measurable outputs (\citeproc{ref-Andersson:2022}{Andersson et al.,
2022}; \citeproc{ref-Miguel:2021}{Miguel et al., 2021};
\citeproc{ref-Wang:2022}{Wang et al., 2022}). A complete
parameterisation often requires a stepwise identification of smaller
sets of parameters from a variety of excitations and different data sets
(\citeproc{ref-Chen:2020}{Chen et al., 2020};
\citeproc{ref-Chu:2019}{Chu et al., 2019}; \citeproc{ref-Kirk:2022}{Kirk
et al., 2023}; \citeproc{ref-Lu:2021}{Lu et al., 2021}).

A generic data fitting optimisation problem may be formulated as:
\begin{equation}
\min_{\mathbf{\theta}} ~ \mathcal{L}_{(\hat{\mathbf{y}_i})}(\mathbf{\theta}) ~~~
\textrm{subject to equations (\ref{dynamics})\textrm{-}(\ref{initial_conditions})}
\label{eqn:parameterisation}
\end{equation}

where \(\mathcal{L} : \mathbf{\theta} \mapsto [0,\infty)\) is a cost
function that quantifies the agreement between the model output
\(\mathbf{y}(t)\) and a sequence of observations
\((\hat{\mathbf{y}_i})\) measured at times \(t_i\). Within the
\texttt{PyBOP} framework, the \texttt{FittingProblem} class packages the
model output along with the measured observations, both of which are
then passed to the cost classes for the computation of the specific cost
function. For gradient-based optimisers, the Jacobian of the cost
function with respect to the unknown parameters,
\((\frac{\partial \mathcal{L}}{\partial \theta})\) is computed for step
size and directional information.

Next, we demonstrate the fitting of synthetic data where the system
parameters are known. In this example problem, we use \texttt{PyBaMM}'s
implementation of the single particle model (SPM) with an added contact
resistance submodel. We assume that the battery model is already
parameterised except for two dynamic parameters, namely the lithium
diffusivity of the negative electrode active material particle (denoted
``negative particle diffusivity'') and the contact resistance. We
generate synthetic data from a one-hour discharge from 100\% state of
charge, to 0\% (denoted as 1C rate), followed by 30 minutes of
relaxation. This data is then corrupted with zero mean Gaussian noise of
amplitude 2mV, shown by the dots in
\autoref{fig:inference-time-landscape} (left). The initial states are
assumed known, although such an assumption is not generally necessary.
The \texttt{PyBOP} repository contains multiple
\href{https://github.com/pybop-team/PyBOP/tree/develop/examples/notebooks}{example
notebooks} that follow a similar inference process for readers that
would like further information. The underlying cost landscape to be
explored by the optimiser is shown in
\autoref{fig:inference-time-landscape} (right) with the initial position
denoted alongside the known system parameters for this synthetic
inference task. In general, the true parameters are not known a priori.

\begin{figure}
\centering
\includegraphics[width=1\textwidth,height=\textheight]{figures/joss/sim-landscape.png}
\caption{The cost landscape for the time-series parameterisation problem
with a root mean squared error cost function.
\label{fig:inference-time-landscape}}
\end{figure}

As mentioned above, \texttt{PyBOP} also provides inference and
optimisation capabilities through numerical computation of the
electrochemical impedance. This is based on the methods presented in the
\texttt{pybamm-eis} package and allows fast impedance computation by
inversion of a sparse mass matrix after scaling and subtraction of the
auto-differentiated Jacobian matrix. More information about this
procedure is available in (\citeproc{ref-pybamm-eis}{Dhoot et al.,
n.d.}). The \autoref{fig:impedance-landscape} below shows the numerical
impedance prediction available in \texttt{PyBOP} alongside the cost
landscape constructed for the inference task. At the time of
publication, gradient-based optimisation and sampling methods are not
available when using an impedance workflow.

\begin{figure}
\centering
\includegraphics[width=1\textwidth,height=\textheight]{figures/joss/impedance.png}
\caption{The cost landscape for the frequency domain impedance
parameterisation problem with a root mean squared error cost function at
5\% SOC. \label{fig:impedance-landscape}}
\end{figure}

To avoid over complicating this example, we will continue with
identification in the time-domain; however, in general these two
simulation methods can be combined for improved system excitation. As
gradient information is available for this problem, the choice of
distance-based cost function and optimiser is not constrained. Due to
the difference in magnitude between the two parameters, we apply the
logarithmic parameter transformation offered by \texttt{PyBOP}. This
transforms the search space of the optimiser to allow for a common step
size between the parameters, which is generally is not required, but
improves convergence in this problem. As a demonstration of the
parameterisation capabilities of \texttt{PyBOP},
\autoref{fig:convergence-min-max} (left) shows the rate of convergence
for each of the distance-minimising cost functions, while
\autoref{fig:convergence-min-max} (right) shows analogous results for
maximising a likelihood. The optimisation is performed with SciPy
Minimize using the gradient-based L-BFGS-B method.

\begin{figure}
\centering
\includegraphics[width=1\textwidth,height=\textheight]{figures/joss/converge.png}
\caption{Convergence in the likelihood functions obtained using various
likelihood functions and the L-BFGS-B algorithm.
\label{fig:convergence-min-max}}
\end{figure}

Next, the performance of the various optimisation algorithms is
presented by category: gradient-based methods in
\autoref{fig:optimiser-inference} (left), evolutionary strategies in
\autoref{fig:optimiser-inference} (middle) and (meta)heuristics in
\autoref{fig:optimiser-inference} (right) for a mean squared error cost
function. Note that the performance of the optimiser depends on the cost
environment, prior information and corresponding hyperparameters for
each specific problem.

\begin{figure}
\centering
\includegraphics[width=1\textwidth,height=\textheight]{figures/joss/optimisers_parameters.png}
\caption{Convergence in the parameter values obtained for the various
optimisation algorithms provided by \texttt{PyBOP}.
\label{fig:optimiser-inference}}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=1\textwidth,height=\textheight]{figures/joss/contour_subplot.png}
\caption{Cost landscape contour plot with corresponding optimisation
traces. The three rows show the gradient-based optimisers (top),
evolution strategies (middle), and (meta)heuristics (bottom).
\label{fig:optimiser-inference}}
\end{figure}

This parameterisation task can also be approached from a Bayesian
perspective, which we will present below using \texttt{PyBOP}`s sampler
methods. The optimisation equation presented in
\autoref{eqn:parameterisation} does not represent the Bayesian parameter
identification task, and as such we introduce Bayes' theorem as,

\begin{equation}
P(\theta|D) = \frac{P(D|\theta)P(\theta)}{P(D)}
\label{eqn:bayes_theorem}
\end{equation}

where, \(P(\theta|D)\) is the posterior, representing the probability
density function of the parameter. \(P(D|\theta)\) is the likelihood
function, which assesses the parameter values alongside a noise model.
\(P(\theta)\) encapsulates the prior knowledge about the parameters, and
finally \(P(D)\) is the model evidence, which acts as a normalising
constant. Our goal in parameter inference is to identify the parameter
values with the highest probability, which can be represented either by
a point-based metric or a posterior. The posterior distribution provides
additional information about the uncertainty of the identified
parameters. Monte Carlo methods are used to sample from the posterior.
The selection of Monte Carlo methods available in \texttt{PyBOP}
includes gradient-based methods such as No-U-Turn
(\citeproc{ref-NUTS:2011}{Hoffman \& Gelman, 2011}) and Hamiltonian
(\citeproc{ref-Hamiltonian:2011}{Brooks et al., 2011}), as well as
heuristic methods such as Differential Evolution
(\citeproc{ref-DiffEvolution:2006}{Braak, 2006}), and finally
conventional methods based on random sampling with rejection criteria
(\citeproc{ref-metropolis:1953}{Metropolis et al., 1953}).
\texttt{PyBOP} offers a sampler class that provides the interface to
these samplers, which are provided by the Probabilistic Inference of
Noise Time-Series (\texttt{PINTS}) package. \autoref{fig:posteriors}
below shows the sampled posterior for the synthetic workflow described
above, using an adaptive covariance-based sampler, Haario Bardenet
(\citeproc{ref-Haario:2001}{Haario et al., 2001}).

\begin{figure}
\centering
\includegraphics[width=1\textwidth,height=\textheight]{figures/joss/posteriors.png}
\caption{Posterior distributions for model parameters alongside
identified noise on the observations. Shaded area denotes confidence
bounds for each parameter. \label{fig:posteriors}}
\end{figure}

\subsection{Design optimisation}\label{design-optimisation}

Design optimisation is supported within \texttt{PyBOP} to guide future
battery design development by identifying parameter sensitivities that
can unlock improvements in battery performance. This problem can be
viewed in a similar way to the parameterisation workflows described
above, but with the aim of maximising a cost function rather than
minimising it. \texttt{PyBOP} performs maximisation by minimising the
negative of the cost function. In design problems, the cost metric is no
longer a distance between two time series, but a metric evaluated on a
model prediction. For example, to maximise the gravimetric energy (or
power) density, the cost is the integral of the discharge energy (or
power) normalised by the cell mass. Such metrics are typically
quantified for operating conditions such as a 1C discharge, at a given
temperature.

In general, design optimisation can be written in the form of a
constrained optimisation problem as: \begin{equation}
\min_{\mathbf{\theta} \in \Omega} ~ \mathcal{L}(\mathbf{\theta}) ~~~
\textrm{subject to equations (\ref{dynamics})\textrm{-}(\ref{initial_conditions})}
\label{eqn:design}
\end{equation}

where \(\mathcal{L} : \mathbf{\theta} \mapsto [0,\infty)\) is a cost
function that quantifies the desirability of the design and \(\Omega\)
is the set of allowable parameter values.

As an example, we consider the problem of maximising the gravimetric
energy density subject to constraints on two of the geometric electrode
parameters (\citeproc{ref-Couto:2023}{Couto et al., 2023}). For this
example, we use the\texttt{PyBaMM} implementation of the single particle
model with electrolyte (SPMe) to investigate the effect of the positive
electrode thickness and the active material volume fraction on the
target cost. Since the active material volume fraction is related to the
electrode porosity, the porosity is defined with a driven constraint
from the volume fraction. In this problem, we estimate the 1C rate from
the theoretical capacity for each iteration of the design. For this
example, we employ the Particle Swarm Optimisation (PSO) algorithm.

\begin{figure}
\centering
\includegraphics[width=1\textwidth,height=\textheight]{figures/joss/design.png}
\caption{The gravimetric energy density landscape alongside the
corresponding initial and optimised voltage profiles for a 1C discharge.
\label{fig:design_gravimetric}}
\end{figure}

\autoref{fig:design_gravimetric} (left) shows the optimiser's search
over the parameter space and (right) the predicted improvement in the
discharge profile between the initial and optimised parameter values,
simulated at their respective 1C rates.

\section{Acknowledgements}\label{acknowledgements}

We gratefully acknowledge all
\href{https://github.com/pybop-team/PyBOP?tab=readme-ov-file\#contributors-}{contributors}
to this package. This work was supported by the Faraday Institution
Multiscale Modelling (MSM) project (ref. FIRG059), UKRI's Horizon Europe
Guarantee (ref. 10038031), and EU IntelLiGent project (ref. 101069765).

\section*{References}\label{references}
\addcontentsline{toc}{section}{References}

\phantomsection\label{refs}
\begin{CSLReferences}{1}{0}
\bibitem[\citeproctext]{ref-Andersson:2022}
Andersson, M., Streb, M., Ko, J. Y., Löfqvist Klass, V., Klett, M.,
Ekström, H., Johansson, M., \& Lindbergh, G. (2022). Parametrization of
physics-based battery models from input-output data: {A} review of
methodology and current research. \emph{Journal of Power Sources},
\emph{521}(November 2021), 230859.
\url{https://doi.org/10.1016/j.jpowsour.2021.230859}

\bibitem[\citeproctext]{ref-DiffEvolution:2006}
Braak, C. J. F. T. (2006). A {Markov} {Chain} {Monte} {Carlo} version of
the genetic algorithm {Differential} {Evolution}: Easy {Bayesian}
computing for real parameter spaces. \emph{Statistics and Computing},
\emph{16}(3), 239--249. \url{https://doi.org/10.1007/s11222-006-8769-1}

\bibitem[\citeproctext]{ref-jax:2018}
Bradbury, J., Frostig, R., Hawkins, P., Johnson, M. J., Leary, C.,
Maclaurin, D., Necula, G., Paszke, A., VanderPlas, J., Wanderman-Milne,
S., \& Zhang, Q. (2018). \emph{{JAX}: Composable transformations of
{P}ython+{N}um{P}y programs} (Version 0.3.13).
\url{http://github.com/jax-ml/jax}

\bibitem[\citeproctext]{ref-Hamiltonian:2011}
Brooks, S., Gelman, A., Jones, G., \& Meng, X.-L. (2011). \emph{Handbook
of markov chain monte carlo}. Chapman; Hall/CRC.
\url{https://doi.org/10.1201/b10905}

\bibitem[\citeproctext]{ref-Planella:2022}
Brosa Planella, F., Ai, W., Boyce, A. M., Ghosh, A., Korotkin, I., Sahu,
S., Sulzer, V., Timms, R., Tranter, T. G., Zyskin, M., Cooper, S. J.,
Edge, J. S., Foster, J. M., Marinescu, M., Wu, B., \& Richardson, G.
(2022). {A Continuum of Physics-Based Lithium-Ion Battery Models
Reviewed}. \emph{Progress in Energy}, \emph{4}(4), 042003.
\url{https://doi.org/10.1088/2516-1083/ac7d31}

\bibitem[\citeproctext]{ref-blackjax:2024}
Cabezas, A., Corenflos, A., Lao, J., \& Louf, R. (2024). \emph{BlackJAX:
Composable {B}ayesian inference in {JAX}}.
\url{https://arxiv.org/abs/2402.10797}

\bibitem[\citeproctext]{ref-Chen:2020}
Chen, C.-H., Brosa Planella, F., O'Regan, K., Gastol, D., Widanage, W.
D., \& Kendrick, E. (2020). Development of experimental techniques for
parameterization of multi-scale lithium-ion battery models.
\emph{Journal of The Electrochemical Society}, \emph{167}(8), 080534.
\url{https://doi.org/10.1149/1945-7111/ab9050}

\bibitem[\citeproctext]{ref-Chu:2019}
Chu, Z., Plett, G. L., Trimboli, M. S., \& Ouyang, M. (2019). A
control-oriented electrochemical model for lithium-ion battery, {P}art
{I}: {L}umped-parameter reduced-order model with constant phase element.
\emph{Journal of Energy Storage}, \emph{25}(August), 100828.
\url{https://doi.org/10.1016/j.est.2019.100828}

\bibitem[\citeproctext]{ref-Clerx:2019}
Clerx, M., Robinson, M., Lambert, B., Lei, C. L., Ghosh, S., Mirams, G.
R., \& Gavaghan, D. J. (2019). Probabilistic inference on noisy time
series ({PINTS}). \emph{Journal of Open Research Software}, \emph{7}(1),
23. \url{https://doi.org/10.5334/jors.252}

\bibitem[\citeproctext]{ref-Couto:2023}
Couto, L. D., Charkhgard, M., Karaman, B., Job, N., \& Kinnaert, M.
(2023). Lithium-ion battery design optimization based on a dimensionless
reduced-order electrochemical model. \emph{Energy}, \emph{263}(PE),
125966. \url{https://doi.org/10.1016/j.energy.2022.125966}

\bibitem[\citeproctext]{ref-optax:2020}
DeepMind, Babuschkin, I., Baumli, K., Bell, A., Bhupatiraju, S., Bruce,
J., Buchlovsky, P., Budden, D., Cai, T., Clark, A., Danihelka, I.,
Dedieu, A., Fantacci, C., Godwin, J., Jones, C., Hemsley, R., Hennigan,
T., Hessel, M., Hou, S., \ldots{} Viola, F. (2020). \emph{The
{D}eep{M}ind {JAX} {E}cosystem}. \url{http://github.com/google-deepmind}

\bibitem[\citeproctext]{ref-pybamm-eis}
Dhoot, R., Timms, R., \& Please, C. (n.d.). \emph{PyBaMM EIS: Efficient
linear algebra methods to determine li-ion battery behaviour} (Version
0.1.4). \url{https://www.github.com/pybamm-team/pybamm-eis}

\bibitem[\citeproctext]{ref-Doyle:1993}
Doyle, M., Fuller, T. F., \& Newman, J. (1993). {Modeling of
Galvanostatic Charge and Discharge of the Lithium/Polymer/Insertion
Cell}. \emph{Journal of The Electrochemical Society}, \emph{140}(6),
1526--1533. \url{https://doi.org/10.1149/1.2221597}

\bibitem[\citeproctext]{ref-Fuller:1994}
Fuller, T. F., Doyle, M., \& Newman, J. (1994). Simulation and
optimization of the dual lithium ion insertion cell. \emph{Journal of
The Electrochemical Society}, \emph{141}(1), 1.
\url{https://doi.org/10.1149/1.2054684}

\bibitem[\citeproctext]{ref-Haario:2001}
Haario, H., Saksman, E., \& Tamminen, J. (2001). An {Adaptive}
{Metropolis} {Algorithm}. \emph{Bernoulli}, \emph{7}(2), 223.
\url{https://doi.org/10.2307/3318737}

\bibitem[\citeproctext]{ref-NUTS:2011}
Hoffman, M. D., \& Gelman, A. (2011). \emph{The no-u-turn sampler:
Adaptively setting path lengths in hamiltonian monte carlo}.
\url{https://arxiv.org/abs/1111.4246}

\bibitem[\citeproctext]{ref-Kirk:2022}
Kirk, T. L., Lewis-Douglas, A., Howey, D., Please, C. P., \& Jon
Chapman, S. (2023). Nonlinear electrochemical impedance spectroscopy for
lithium-ion battery model parameterization. \emph{Journal of The
Electrochemical Society}, \emph{170}(1), 010514.
\url{https://doi.org/10.1149/1945-7111/acada7}

\bibitem[\citeproctext]{ref-BPX:2023}
Korotkin, I., Timms, R., Foster, J. F., Dickinson, E., \& Robinson, M.
(2023). Battery parameter eXchange. In \emph{GitHub repository}. The
Faraday Institution. \url{https://github.com/FaradayInstitution/BPX}

\bibitem[\citeproctext]{ref-Lu:2021}
Lu, D., Scott Trimboli, M., Fan, G., Zhang, R., \& Plett, G. L. (2021).
Implementation of a physics-based model for half-cell open-circuit
potential and full-cell open-circuit voltage estimates: {P}art {II}.
Processing full-cell data. \emph{Journal of The Electrochemical
Society}, \emph{168}(7), 070533.
\url{https://doi.org/10.1149/1945-7111/ac11a5}

\bibitem[\citeproctext]{ref-metropolis:1953}
Metropolis, N., Rosenbluth, A. W., Rosenbluth, M. N., Teller, A. H., \&
Teller, E. (1953). Equation of {State} {Calculations} by {Fast}
{Computing} {Machines}. \emph{The Journal of Chemical Physics},
\emph{21}(6), 1087--1092. \url{https://doi.org/10.1063/1.1699114}

\bibitem[\citeproctext]{ref-Miguel:2021}
Miguel, E., Plett, G. L., Trimboli, M. S., Oca, L., Iraola, U., \&
Bekaert, E. (2021). Review of computational parameter estimation methods
for electrochemical models. \emph{Journal of Energy Storage},
\emph{44}(PB), 103388. \url{https://doi.org/10.1016/j.est.2021.103388}

\bibitem[\citeproctext]{ref-numpyro:2019}
Phan, D., Pradhan, N., \& Jankowiak, M. (2019). Composable effects for
flexible and accelerated probabilistic programming in NumPyro.
\emph{arXiv Preprint arXiv:1912.11554}.

\bibitem[\citeproctext]{ref-Sulzer:2021}
Sulzer, V., Marquis, S. G., Timms, R., Robinson, M., \& Chapman, S. J.
(2021). {Python Battery Mathematical Modelling (PyBaMM)}. \emph{Journal
of Open Research Software}, \emph{9}(1), 14.
\url{https://doi.org/10.5334/jors.309}

\bibitem[\citeproctext]{ref-Tranter2022}
Tranter, T. G., Timms, R., Sulzer, V., Planella, F. B., Wiggins, G. M.,
Karra, S. V., Agarwal, P., Chopra, S., Allu, S., Shearing, P. R., \&
Brett, D. J. l. (2022). Liionpack: A python package for simulating packs
of batteries with PyBaMM. \emph{Journal of Open Source Software},
\emph{7}(70), 4051. \url{https://doi.org/10.21105/joss.04051}

\bibitem[\citeproctext]{ref-Verbrugge:2017}
Verbrugge, M., Baker, D., Koch, B., Xiao, X., \& Gu, W. (2017).
Thermodynamic model for substitutional materials: Application to
lithiated graphite, spinel manganese oxide, iron phosphate, and layered
nickel-manganese-cobalt oxide. \emph{Journal of The Electrochemical
Society}, \emph{164}(11), E3243.
\url{https://doi.org/10.1149/2.0341708jes}

\bibitem[\citeproctext]{ref-SciPy:2020}
Virtanen, P., Gommers, R., Oliphant, T. E., Haberland, M., Reddy, T.,
Cournapeau, D., Burovski, E., Peterson, P., Weckesser, W., Bright, J.,
van der Walt, S. J., Brett, M., Wilson, J., Millman, K. J., Mayorov, N.,
Nelson, A. R. J., Jones, E., Kern, R., Larson, E., \ldots{} SciPy 1.0
Contributors. (2020). {{SciPy} 1.0: Fundamental Algorithms for
Scientific Computing in Python}. \emph{Nature Methods}, \emph{17},
261--272. \url{https://doi.org/10.1038/s41592-019-0686-2}

\bibitem[\citeproctext]{ref-Wang:2022}
Wang, A. A., O'Kane, S. E. J., Brosa Planella, F., Houx, J. L., O'Regan,
K., Zyskin, M., Edge, J., Monroe, C. W., Cooper, S. J., Howey, D. A.,
Kendrick, E., \& Foster, J. M. (2022). Review of parameterisation and a
novel database {(LiionDB)} for continuum {Li-ion} battery models.
\emph{Progress in Energy}, \emph{4}(3), 032004.
\url{https://doi.org/10.1088/2516-1083/ac692c}

\end{CSLReferences}

\end{document}